---
title: "Data Sci Boot Camp"
author: "Germán Silva"
date: "7/13/2022"
output: html_document
---

# Overview

Exploring, analysizing, and visualizing *Artemisia* data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE) # shows code in final output, but hides messages and warnings

# Attach necessary libraries

library(tidyverse)
library(here)
library(kableExtra)
library(ggbeeswarm)
```


## Reading and Summarizing the data:
```{r}
bulk_den <- read_csv(here("NCOS_Bulk_Den_Samples.csv")) %>%  # reads in the csv so we can work with it in the code
  drop_na() # drops all rows with an NA

bulk_den_average <- bulk_den %>% #starts new table
  group_by(Zone, `GPS_point_#`) %>% # tells R to group by these two identifying variables
  summarize(mean_bulk = mean(bulk_density_g_cm_3)) # returns a table with grouping variables and the stat we requested

bulk_den_sum <- bulk_den_average %>% # calls the averaged data
  group_by(Zone) %>% #groups by zone only
  summarize(min = min(mean_bulk), # returns minimum value
            max = max(mean_bulk), # returns maximum value
            sd = sd(mean_bulk), # returns group's standard deviation
            mean = mean(mean_bulk), # returns group's mean
            sample_size = n()) # returns sample size

kable(bulk_den_sum, # starts the creation of a table with the data from the object called
      col.names = c("Zone",
                    "Min. Bulk Density (g/cm^3)",
                    "Max. Bulk Denisty (g/cm^3)",
                    "Standard Deviation (g/cm^3)",
                    "Mean Bulk Density (g/cm^3)",
                    "Sample Size")) %>%  # renames all the column names to professional looking titles
  kable_classic(full_width = FALSE) # themes the table


```

```{r}
ggplot(bulk_den_average, aes(x = Zone, y = mean_bulk, color = Zone))+ 
  geom_beeswarm(aes(fill=Zone), shape= 21, color = "black")+
  geom_boxplot(aes(fill=Zone),
               color = "black",
               alpha=0.45,
               width = 0.2,
               outlier.color = NA)+
   scale_fill_manual(values = c("#DEADA1", "#6d748c"))+
  ggtitle("Bulk Density Across Zones")+
  labs(x= "Zone",
       y = "Bulk Density (g/cm3)")+
   theme(legend.position = "none",
      plot.title = element_text(color = "#5b4f41", size = 16, hjust = 0.45),
        plot.background = element_rect("white"),
        panel.background = element_rect("#f8f8f8"),
        panel.grid = element_line(linetype= "longdash", color = "#f0ece1"),
        axis.text = element_text(color = "#5b4f41", size = 7),
        axis.title = element_text(color = "#5b4f41"),
        strip.background = element_rect("white"),
        axis.line = element_line(color = "#5b4f41"),
        plot.caption = element_text(size = 9, hjust = -0.1))
```


## Hypothesis testing

Questions:

- Is there a significant difference in mean soil bulk density between the two zones? Is there a significant difference between longest branch length in the two zones?
- Does the data fit the assumptions of parametric significance testing?

### Exploring the data

Histogram:
```{r}
ggplot(data = bulk_den_average, aes(x = mean_bulk))+ # starts creation of graph, aes() refers to aesthetics and decides how things are displayed from the variables
  geom_histogram()+ # turns graph into a histogram
  facet_wrap(~Zone) # makes two panels based on Zone grouping

ggplot(bulk_den_average, aes(x = mean_bulk, color = Zone, fill = Zone))+
  geom_histogram(alpha = 0.5)
```


QQ plot:
```{r}
ggplot(bulk_den_average, aes(sample = mean_bulk))+ # starts graph with mean bulk density as the sample/graphing data
  geom_qq()+ # makes a QQ plot, a exploratory/diagnostic plot
  facet_wrap(~Zone) # panels by Zone Grouping
```

### T-test

```{r}
f1 <- bulk_den_average %>% 
  filter(Zone == "F1")

f3 <- bulk_den_average %>% 
  filter(Zone == "F3")

cohen_d <- effsize::cohen.d(f1$mean_bulk, f3$mean_bulk)

cohen_d

t_test <- t.test(f1$mean_bulk, f3$mean_bulk)

t_test
```

