---
title: "Data Sci Boot Camp"
author: "Germán Silva"
date: "7/13/2022"
output: html_document
---

# Overview

Exploring, analyzing, and visualizing *Artemisia* data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE) # shows code in final output, but hides messages and warnings

# Attach necessary libraries

# significance tests and data exploration
library(tidyverse)
library(here)
library(kableExtra)
library(ggbeeswarm)

# spatial analysis
library(sf)
library(tmap)
library(gstat)
library(stars)
library(SpatialKDE)
```


# Reading and Summarizing the data:

## Reading and cleaning data

### Bulk Density
```{r}
# pipe operator: Windows (ctrl+shift+m), MAC (cmd+shift+m)

bulk_den <- read_csv(here("NCOS_Bulk_Den_Samples.csv")) %>%  # reads in the csv so we can work with it in the code
  drop_na() # drops all rows with an NA

bulk_den_average <- bulk_den %>% #starts new table
  janitor::clean_names() %>% 
  group_by(zone, gps_point_number) %>% # tells R to group by these two identifying variables
  summarize(mean_bulk = mean(bulk_density_g_cm_3)) # returns a table with grouping variables and the stat we requested

bulk_den_sum <- bulk_den_average %>% # calls the averaged data
  group_by(zone) %>% #groups by zone only
  summarize(min = min(mean_bulk), # returns minimum value
            max = max(mean_bulk), # returns maximum value
            sd = sd(mean_bulk), # returns group's standard deviation
            mean = mean(mean_bulk), # returns group's mean
            sample_size = n()) # returns sample size

kable(bulk_den_sum, # starts the creation of a table with the data from the object called
      col.names = c("Zone",
                    "Min. Bulk Density (g/cm^3)",
                    "Max. Bulk Denisty (g/cm^3)",
                    "Standard Deviation (g/cm^3)",
                    "Mean Bulk Density (g/cm^3)",
                    "Sample Size")) %>%  # renames all the column names to professional looking titles
  kable_classic(full_width = FALSE) # themes the table


```

### Artemisia

```{r}
art_cal <- read_csv(here("artemisia_transects.csv")) %>% 
  janitor::clean_names()

art_cal_sum <- art_cal %>% 
  group_by(zone) %>% #groups by zone only
  summarize(min = min(longest_branch_cm), # returns minimum value
            max = max(longest_branch_cm), # returns maximum value
            sd = sd(longest_branch_cm), # returns group's standard deviation
            mean = mean(longest_branch_cm), # returns group's mean
            sample_size = n())

kable(art_cal_sum, # starts the creation of a table with the data from the object called
      col.names = c("Zone",
                    "Min. Length (cm)",
                    "Max. Length (cm)",
                    "Standard Deviation (cm)",
                    "Mean Length(cm)",
                    "Sample Size")) %>%  # renames all the column names to professional looking titles
  kable_classic(full_width = FALSE) # themes the table

```


## Visualizing

### Bulk Density
```{r}
ggplot(bulk_den_average, aes(x = zone, y = mean_bulk, color = zone))+ 
  geom_beeswarm(aes(fill=zone), shape= 21, color = "black")+
  geom_boxplot(aes(fill=zone),
               color = "black",
               alpha=0.45,
               width = 0.2,
               outlier.color = NA)
```

### Artemisia

```{r}
ggplot(art_cal, aes(x = zone, y = longest_branch_cm, color = zone))+ 
  geom_beeswarm(aes(fill=zone), shape= 21, color = "black")+
  geom_boxplot(aes(fill=zone),
               color = "black",
               alpha=0.45,
               width = 0.2,
               outlier.color = NA)
```


# Hypothesis testing

Questions:

- Is there a significant difference in mean soil bulk density between the two zones? Is there a significant difference between longest branch length in the two zones?
- Does the data fit the assumptions of parametric significance testing?

## Exploring the data

### Soil Bulk Density

Histogram:
```{r}
ggplot(data = bulk_den_average, aes(x = mean_bulk))+ # starts creation of graph, aes() refers to aesthetics and decides how things are displayed from the variables
  geom_histogram()+ # turns graph into a histogram
  facet_wrap(~zone) # makes two panels based on Zone grouping

ggplot(bulk_den_average, aes(x = mean_bulk, color = zone, fill = zone))+
  geom_histogram(alpha = 0.5)
```


QQ plot:
```{r}
ggplot(bulk_den_average, aes(sample = mean_bulk))+ # starts graph with mean bulk density as the sample/graphing data
  geom_qq()+ # makes a QQ plot, a exploratory/diagnostic plot
  facet_wrap(~zone) # panels by Zone Grouping
```


### Artemisia

Histogram:
```{r}
ggplot(data = art_cal, aes(x = longest_branch_cm))+ # starts creation of graph, aes() refers to aesthetics and decides how things are displayed from the variables
  geom_histogram()+ # turns graph into a histogram
  facet_wrap(~zone) # makes two panels based on Zone grouping

ggplot(art_cal, aes(x = longest_branch_cm, color = zone, fill = zone))+
  geom_histogram(alpha = 0.5)
```

QQ plot:
```{r}
ggplot(art_cal, aes(sample = longest_branch_cm))+ # starts graph with mean bulk density as the sample/graphing data
  geom_qq()+ # makes a QQ plot, a exploratory/diagnostic plot
  facet_wrap(~zone) # panels by Zone Grouping
```

## T-test

### Bulk Density

```{r}
f1 <- bulk_den_average %>% 
  filter(zone == "F1")

f3 <- bulk_den_average %>% 
  filter(zone == "F3")

cohen_d <- effsize::cohen.d(f1$mean_bulk, f3$mean_bulk)

cohen_d

t_test <- t.test(f1$mean_bulk, f3$mean_bulk)

t_test
```

### Artemisia
```{r}
f1_art <- art_cal %>% 
  filter(zone == "F1")

f3_art <- art_cal %>% 
  filter(zone == "F3")

cohen_d_art <- effsize::cohen.d(f1_art$longest_branch_cm, f3_art$longest_branch_cm)

cohen_d_art

t_test_art <- t.test(f1_art$longest_branch_cm, f3_art$longest_branch_cm)

t_test_art
```

Interpret these results with the group


# Spatial Interpolation

## Read shapefiles and set coordinates

```{r}
soil_sf <- read_sf(here("gis_data", "bulk_density.shp")) %>%
  janitor::clean_names() %>% 
  mutate(zone = case_when(
    cid == 0 ~ "F1",
    cid == 1 ~ "F3"),
    zone_sample = objectid) %>% 
  select(sample_id, zone, zone_sample, bulk_den)

zone_sf <- read_sf(here("gis_data", "zones.shp")) %>% 
  select(Zone)

art_sf <- read_sf(here("gis_data", "all_artemisia_correct.shp"))


# change coordinate system to match others

zone_sf <- st_transform(zone_sf, st_crs(soil_sf))
```

## Example Maps

```{r}
ggplot() +
  geom_sf(data = zone_sf)+
  geom_sf(data = art_sf, aes(color = branch_cm))
```

```{r}
tmap_mode(mode = "view")

tm_shape(zone_sf)+
  tm_polygons()+
  tm_shape(art_sf)+
  tm_dots()
```

## Clip data

```{r}
soil_clip <- soil_sf %>% 
  st_intersection(zone_sf)

art_clip <- art_sf %>% 
  st_intersection(zone_sf)
```


## Create a variogram and variogram model

```{r}
soil_vgm <- gstat::variogram(bulk_den ~ 1,
                                data = soil_clip)

plot(soil_vgm)
```

```{r}
soil_vgm_fit <- fit.variogram(soil_vgm, 
                                 model = vgm(model = "Exp",
                                             nugget = 10,
                                             sill = 1000, 
                                             range = 0.01))

plot(soil_vgm, soil_vgm_fit)
```


## Interpolation by Kriging

```{r}
zone_grid <- st_bbox(zone_sf) %>% 
  st_as_stars() %>% 
  st_set_crs(st_crs(zone_sf)) %>% 
  st_crop(zone_sf)
  

soil_krige <- krige(bulk_den ~ 1, 
                       soil_clip,
                       zone_grid,
                       model = soil_vgm_fit)
plot(soil_krige)

krige_df <- as.data.frame(soil_krige) %>% 
  st_as_sf(coords = c("x", "y")) %>% 
  drop_na(var1.pred) %>% 
  st_set_crs(st_crs(art_clip))

ggplot()+
  geom_sf(data = krige_df,
          aes(color = var1.pred)) +
  scale_color_gradient(low = "blue", high = "yellow")+
  geom_sf(data = art_clip)
```

## Spatial Join

```{r}
art_soil <-  art_clip %>%
  st_join(krige_df, st_is_within_distance, dist = .5) %>% 
  group_by(ID, CID) %>% 
  summarize(branch_cm = mean(branch_cm),
            bulk_den = mean(var1.pred)) %>% 
  ungroup()

```


## Extract slope and aspect

```{r}

dem <- read_stars(here("gis_data", "dem.tif")) %>% 
  st_extract(art_soil)

slope <- read_stars(here("gis_data", "slope.tif")) %>% 
  st_extract(art_soil)

aspect <- read_stars(here("gis_data", "aspect.tif")) %>% 
  st_extract(art_soil)

art_soil <- art_soil %>% 
  st_join(y = dem) %>% 
  st_join(y = slope) %>% 
  st_join(y = aspect)
```


# Regression and Pattern Visualization

## Visualization

```{r}
#visualize data
ggplot(art_soil, aes(x = bulk_den, y = branch_cm))+
  geom_point()+
  geom_smooth(method = "lm")
```


## Linear model (OLS)

```{r}
# create
art_soil_lm <- lm(branch_cm ~ bulk_den, art_soil)

art_soil_lm

summary(art_soil_lm)

#diagnostics 
plot(art_soil_lm)
```


## Multiple Linear Regression

```{r}
art_mlr <- glm(branch_cm ~ bulk_den + dem.tif + slope.tif + aspect.tif,
               data = art_soil)

art_mlr

summary(art_mlr)

plot(art_mlr)

```


```{r}
art_plsr <- caret::train(branch_cm ~ bulk_den + dem.tif + slope.tif + aspect.tif,
                         data = art_soil,
                         method = "pls",
                         trControl = trainControl(method = "cv", number = 10))
```



## Correlation

```{r}
art_cor <- cor.test(art_soil$branch_cm, art_soil$bulk_den)

art_cor
```


## Optional: Kernel Density and Density map
```{r}
art_points <- art_soil %>% 
  st_as_sf(coords = c("x", "y"), dim= "xy") %>% 
  st_set_crs(st_crs(art_clip)) %>% 
  st_transform(crs = 26910) %>% 
  select()

zone_utm <- zone_sf %>% 
  st_transform(crs = 26910)

cell <- 1
band <- 1.5

grid_art <- art_points %>% 
  create_grid_rectangular(cell_size = cell, side_offset = band)

kde <- art_points %>% 
  kde(band_width = band, kernel = "quartic", grid = grid_art) %>% 
  st_intersection(zone_utm) %>% 
  filter(kde_value != 0)

kde_sum <- kde %>% 
  group_by(Zone) %>% 
  summarize(mean_den = mean(kde_value))

ggplot(kde, aes(fill= kde_value))+
  geom_sf()
  
```

